---
title: "FARS"
author: "Marco Mello"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FARS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(FARS)
```

This is the final Peer-graded Assignment from Coursera's Building an R Package. 

The `FARS` package contains functions provided for the students. These functions were written to analyze data from the [US National Highway Traffic Safety Administration's Fatality Analysis Reporting System](https://www.nhtsa.gov) (FARS). FARS is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.


## Data source

Unfortunately, since the course was last updated, the FARS website changed considerably and data are more difficult to find now. Therefore, the data files used to test the functions included in this package were downloaded directly from the [course's website](https://d3c33hcgiwev3.cloudfront.net/_e1adac2a5f05192dc8780f3944feec13_fars_data.zip?Expires=1604534400&Signature=RTTvUAKz8DKWFiSEcAetz-uqBOzv79moF3uW0TbPA8k8fvG6Pa~o1wndjQsxlNKDxFWVONipYXks9nQ~G3zuKUfDj0ufW2d0ITHZlmi9VqjXjoTbW4MjXzoLX8I-CLjs-OG4VzhK6dewCktT1k2mIVfRodd72Kzu3BQTcqK5lZI_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A).

Three data files from FARS dowloaded from the course's website were used to test this package:

* accident_2013.csv.bz2

* accident_2014.csv.bz2

* accident_2015.csv.bz2

If you want to download equivalent data files from other years, visit the [Download Your Data](https://cdan.nhtsa.gov) section of the FARS website.


## Packages

You need to load the following dependencies before starting:

```{r, echo=TRUE, include=TRUE}
library(dplyr)
library(maps)
library(readr)
library(tidyr)
```


## Functions

The package contains five functions used to read, name, filter, summarize, and plot the data.

1. `fars_read`: This function reads data in CSV format into R. First, it checks whether the file already exists in the working directory. If the file does not exist, you get an error message.Second, if the CSV file exists, it is read into R. Third, the data are converted to a tibble.

2. `make_filename`: This function makes a file name. First, the inputted year will be converted to an integer. The value must be a simple year without quotes. Second, the year will be added to the middle of a string that represents the main pattern of the file name, in the placeholder for an integer. 

3. `fars_read_years`: This function reads specific variables from the data. It reads two specific columns, MONTH and year, from the input files, and then stores them as data frames in a list. It works by calling the function `make_filename` from within.

4. `fars_summarize_years`: This function summarizes the accidents by year and month for the data frames imported. It calls the function `fars_read_years` from within.

5. `fars_map_state`: This function plots a map with the locations of accidents for the selected state and year.It calls the functions `make_filename` and `fars_read` from within. 
